---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

# concise

<!-- badges: start -->
<!-- badges: end -->

## Overview
`concise` functions are designed to make clean, intelligible lambda functions
to keep your code concise. They are modelled on common `tidyverse` functions like
`purrr::map` and `dplyr::mutate`, but with a layer of syntactic sugar to make
anonymous functions that condense a paragraph's worth of code into a single line.

* Refer to data columns directly -- avoid placeholder pronouns like `.x` or `..1`
and instead refer to your data by name.
* Helpful shorthand gives access to vector properties like row number, vector length
and the vector names inside the iterative function.
* Interact with and preserve groups created by `dplyr::group_by`.
* Access the entire data column as well as individual elements to write custom
summary and window functions such as moving averages.
* Refer to anonymous functions inside their own definitions to write recursive
functions.

## Installation

You can install the development version of concise from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("domjarkey/concise")
```

## Usage

### `cmutate`

`cmutate` performs the same role as `dplyr::mutate` but with the option of evaluating
column definitions as iterative lambda functions.

This allows for the easy use of non-vectorised functions.

```{r cmutate, message=FALSE}
library(concise)
library(dplyr)
library(purrr)

tibble(value = list("a", "b", NULL, "d", NULL)) |>
    cmutate(value_exists ~ !is.null(value))
```

This can side-step the problem of having to use mapping functions that obfuscate
the intent of your code. Consider the problem of picking the largest element from
amongst three columns. Naively, we might try a simple application of the `max`
function:

```{r}
df <- tibble(
    x = c(29L, 11L, 72L, 81L, 27L, 61L, 42L, 26L, 57L, 39L),
    y = c(38L, 80L, 98L, 93L, 34L, 26L, 4L, 31L, 18L, 69L),
    z = c(31L, 83L, 91L, 69L, 82L, 65L, 75L, 3L, 20L, 71L)
)

df |> mutate(
    naive_largest = max(x, y, z)
)
```

Clearly this fails because `max` lumps the values of all the rows in `x`, `y`,
and `z` instead of evaluating the function row by row. We might use `dplyr::rowwise`
to overcome this problem, but this is an undesirable alternative for a couple of 
reasons. First, `rowwise` is very slow, especially for larger data frames. Second, 
`rowwise` effectively works by grouping every row into its own separate group,
meaning existing groups are lost, and the data may need to be ungrouped afterwards
if further mutations need to be performed.

```{r}
# More like slowwise
df |> 
    rowwise() |>
    mutate(rowwise_largest = max(x, y, z)) |>
    ungroup()
```

The better way to perform this operation that requires row-by-row computation
taking multiple data columns as inputs is with the `purrr::pmap` function. `pmap`
can either take a formula to specify the expression of the lambda function,
referring to each input column by its position, or (in more recent versions of R)
an anonymous function defined using `\\\\(x, y, z) expr` notation. These look like
this:

```{r}
df |>
    mutate(
        pmap_largest_v1 = pmap_int(list(x, y, z), ~ max(..1, ..2, ..3)),
        pmap_largest_v2 = pmap_int(list(x, y, z), \(x, y, z) max(x, y, z))
    )
```

Of course these work a lot faster than `rowwise` (the second version is fastest),
but what they gain in speed they lose in brevity. Both end up being much longer
than might initially be expected for a relatively simple operation, and the first
is particularly hard to understand at a glance as the `..x` pronouns obscure which
variables they are actually referring to. The purpose of the `cmutate` function is
to avoid this trade-off between speed and intelligibility, as can be seen below:

```{r}
df |> cmutate(concise_largest ~ max(x, y, z))
```


### `rmap`

`rmap` works similarly to `purrr::pmap` except the input data frame does not need
to be subset to only those columns used in the function, and the data columns can
be directly referred to in the anonymous function.
