---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

# concise

<!-- badges: start -->
<!-- badges: end -->

## Overview
`concise` functions are designed to make clean, intelligible lambda functions
to keep your code concise. They are modelled on common `tidyverse` functions like
`purrr::map` and `dplyr::mutate`, but with a layer of syntactic sugar to make
anonymous functions that condense a paragraph's worth of code into a single line.

* Refer to data columns directly -- avoid placeholder pronouns like `.x` or `..1`
and instead refer to your data by name.
* Helpful shorthand gives access to vector properties like row number, vector length
and the vector names inside the iterative function.
* Interact with and preserve groups created by `dplyr::group_by`.
* Access the entire data column as well as individual elements to write custom
summary and window functions such as moving averages.
* Refer to anonymous functions inside their own definitions to write recursive
functions.

## Installation

You can install the development version of concise from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("domjarkey/concise")
```

## Usage

### `cmutate`

`cmutate` performs the same role as `dplyr::mutate` but with the option of evaluating
column definitions as iterative lambda functions.

This allows for the easy application of non-vectorised functions on a row-by-row
basis.

```{r cmutate, message=FALSE}
library(concise)
library(dplyr)
library(purrr)

tibble(value = list("a", "b", NULL, "d", NULL)) |>
    cmutate(value_exists ~ !is.null(value))
```

#### Examples

Find the largest value out of three columns...

```{r}
df <- tibble::tibble(
    x = c(29L, 11L, 72L, 81L, 27L, 61L, 42L, 26L, 57L, 39L),
    y = c(38L, 80L, 98L, 93L, 34L, 26L, 4L, 31L, 18L, 69L),
    z = c(31L, 83L, 91L, 69L, 82L, 65L, 75L, 3L, 20L, 71L)
)
```

:::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px;"}

::: {}

`concise` syntax:

```{r}
library(concise)
df |> cmutate(
    largest ~ max(x, y, z)
)
```
:::

::: {}

Equivalent code:

```{r message=FALSE}
library(tidyverse)
df |> mutate(
    largest = pmap_int(list(x, y, z), \(x, y, z) max(x, y, z))
)
```
:::

::::

### `rmap`

`rmap` works similarly to `purrr::pmap` except the input data frame does not need
to be subset to only those columns used in the function, and the data columns can
be directly referred to in the anonymous function.
